---
title: Analysis of traffic at the intersections around Fullarton and Glen Osmond Roads
author: "Ben Moretti"
date: "07/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(janitor)
library(tidyquant)
library(scales)
library(fs)
library(kableExtra)

```

# Introduction

This notebook analyses some data from the intersections around Fullarton / Glen Osmond Road in Adelaide, South Australia. 

(put a map here)

# Data

Data is derived from the South Australian Government data portal 

https://data.sa.gov.au/data/dataset/traffic-volumes-on-top-40-intersections-in-sa

"A ranking of the top 40 intersections based on the number of vehicles passing through each intersection."

Data were downloaded in CSV format to a subdirectory, as follows 

```{r}
fs::dir_ls("../data/")
```

Each file has the following metadata as defined 

* Rank - Relative importance
* Location - Description of the intersecting roads 
* Vehicle Exposure - "A measure of traffic volumes passing through intersections. It is calculated by adding the estimated Annual Average Daily Traffic (est. AADT) volumes on all arms of the intersection and dividing the result by two. Vehicle Exposure estimates are based on data obtained from the most recent single day 11 hour manual turning count surveys conducted at the intersections. 


# Analysis



## Data Gathering

First, the data are read from CSV and have the recorded year appended 

```{r}


# All of the data have been downloaded to the data directory from the repository above 

# 2017 data
data_2017_tbl <- read_csv(file = "../data/top-40-at-grade-arterial-road-intersections-for-all-vehicles-2017.csv", skip=2) %>%
  clean_names() %>%
  mutate(
    recorded_year = 2017
  )

# 2018 data 
data_2018_tbl <- read_csv(file = "../data/top-40-at-grade-arterial-road-intersections-for-all-vehicles-2018.csv", skip=2) %>%
  clean_names() %>%
  mutate(
    recorded_year = 2018
  )

# 2019 data
data_2019_tbl <- read_csv(file = "../data/top-40-at-grade-arterial-road-intersections-for-all-vehicles-2019.csv", skip=2) %>%
  clean_names() %>%
  mutate(
    recorded_year = 2019
  )

# 2020 data 
data_2020_tbl <- read_csv(file = "../data/top-40-at-grade-arterial-road-intersections-for-all-vehicles-2020.csv", skip=2) %>%
  clean_names() %>%
  mutate(
    recorded_year = 2020
  )


```

## Tidy 

Once read the data are combined into a single data frame

```{r}
# combine the data into a single data frame
data_combined_tbl <- bind_rows(data_2017_tbl,
                               data_2018_tbl,
                               data_2019_tbl,
                               data_2020_tbl)

```


## Explore

The data structure can be reviewed here  

```{r}
# data frame structure
data_combined_tbl %>% glimpse()

```

Look at the distinct names for the locations

```{r}
data_combined_tbl %>%
  distinct(location) %>%
  arrange(location) %>%
  kbl() %>%
  kable_styling()
```

Looks like we just need the following locations 

```{r}
data_combined_tbl %>%  
  filter(str_detect(location, "GLEN OSMOND") | str_detect(location, "FULLARTON") | str_detect(location, "UNLEY") | str_detect(location, "PORTRUSH"))  %>%
  filter(! str_detect(location, "PAYNEHAM")) %>%
  distinct(location) 
```



Assign the locations we want to a list 

```{r}
location_lst <- data_combined_tbl %>%  
  filter(str_detect(location, "GLEN OSMOND") | str_detect(location, "FULLARTON") | str_detect(location, "UNLEY") | str_detect(location, "PORTRUSH"))  %>%
  filter(! str_detect(location, "PAYNEHAM")) %>%
  distinct(location) %>%
  pull()

```

# Visualise 

Plot out the data in time series 

```{r}
intersections_plot <- data_combined_tbl %>%  
  
  # convert values into thousands
  mutate(
    vehicle_exposure = vehicle_exposure / 1000
  ) %>%
  
  # only get the required locations
  filter(location %in% location_lst) %>% 
  
  # clean up location so that it is a bit tidier and shorter 
  mutate(
    location_text = str_remove_all(location, " ROAD"),
    location_text = str_trim(location, side="both")
  ) %>%
  
  # plot a line plot with points for the recorded year vs vehicle exposure with colour by location and a label of the number
  ggplot(aes(x=recorded_year, y=vehicle_exposure, colour=location_text, label=vehicle_exposure)) +
  geom_line(size=2) +
  geom_point(size=3) +
  geom_text(aes(label=vehicle_exposure),hjust=0, vjust=2) +
  
  # facet wrap by location
  facet_wrap(vars(location_text)) +
  ylim(0,80) +
  
  # theme stuff
  theme_tq() +
  scale_colour_tq() +
  
  # informative labels
  labs(
    title="Intersection Traffic Volumes in or around Glen Osmond Road",
    subtitle = str_glue("Derived from \"Traffic volumes on top 40 intersections in SA\". 
    
                         Source: https://data.sa.gov.au/data/dataset/traffic-volumes-on-top-40-intersections-in-sa
                         
                         Vehicle Exposure: A measure of traffic volumes passing through intersections. It is calculated by adding the estimated Annual Average Daily Traffic (est. AADT) volumes 
                         on all arms of the intersection and dividing the result by two. Vehicle Exposure estimates are based on data obtained from the most recent single day 11 hour manual 
                         turning count surveys conducted at the intersections."),
    x = "Recorded Year",
    y = "Vehicle Exposure ('000)",
    caption = "4 June 2021, Ben Moretti @morebento",
    colour = "Location"
  ) 

```

```{r fig.height=7, fig.width=10}
intersections_plot
```
Save it as a plot 

```{r}
ggsave(
  file = "../plots/glen_osmond_road_intersection_traffic.png", 
  plot = intersections_plot, 
  width = 297, 
  height = 210, 
  units = "mm"
)
```